---
title: "eda"
author: "ShufanXia"
date: "10/12/2021"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(lme4)
library(stargazer)
library(lattice)
library(caret)
library(influence.ME)
```

```{r}
load("./Data/streetrx.RData")
head(streetrx)
colnames(streetrx)
streetrx_morphine <- streetrx[streetrx['api_temp']=="morphine",]
nrow(streetrx_morphine)
streetrx_morphine<-na.omit(streetrx_morphine)
streetrx_morphine <- streetrx_morphine[,! colnames(streetrx_morphine) %in% c("yq_pdate","price_date","city","country","Primary_Reason")] #drop columns
#drop entry row with = ""
streetrx_morphine$source <- ifelse(streetrx_morphine$source=="", "Unknown",as.character(streetrx_morphine$source))
#rename state = USA to unkown
streetrx_morphine$state <- ifelse(streetrx_morphine$state=="USA","Unknown",as.character(streetrx_morphine$state))
streetrx_morphine<-streetrx_morphine %>%filter(streetrx_morphine$state!="")

nrow(streetrx_morphine)
head(streetrx_morphine)
```
```{r}
streetrx_morphine$logppm<-log(streetrx_morphine$ppm) # add a log column
streetrx_morphine$mgstr_c <- scale(streetrx_morphine$mgstr) #center mgstr
#factorize categorical variable: source, form_temp and bulk_purchase
streetrx_morphine$bulk_fac <- factor(streetrx_morphine$bulk_purchase,levels=c("0 Not bulk purchase","1 Bulk purchase"),labels=c("No","Yes"))
streetrx_morphine$source_rename <- ifelse(streetrx_morphine$source%in%c("Personal","Heard it","Internet","Internet Pharmacy","Unknown"),as.character(streetrx_morphine$source),"Internet")
streetrx_morphine$source_fac <- factor(streetrx_morphine$source_rename)
streetrx_morphine$form_fac <- factor(streetrx_morphine$form_temp)
levels(streetrx_morphine$form_fac)

```
```{r}
## explore the response variable, price per milogram
summary(streetrx_morphine$ppm)
ggplot(data=streetrx_morphine,aes(x=streetrx_morphine$ppm))+geom_histogram(bins=50)
ggplot(data=streetrx_morphine,aes(x=streetrx_morphine$ppm))+geom_boxplot()
qqnorm(streetrx_morphine$ppm)
# super skewed, it is not even outliers are changing it
```

```{r}
#try some transformation on ppm

#better, add log_ppm, one entry has ppm=0.drop it
streetrx_morphine <- streetrx_morphine[streetrx_morphine$ppm!=0,]
ggplot(data=streetrx_morphine,aes(x=streetrx_morphine$logppm))+geom_histogram(bins=50)
ggplot(data=streetrx_morphine,aes(x=streetrx_morphine$logppm))+geom_boxplot()
qqnorm(streetrx_morphine$logppm)
###??? much better, sqrt does not help at all
```

```{r}
#by region?
streetrx_morphine %>% count(USA_region)

ggplot(streetrx_morphine,
aes(x=USA_region, y=logppm,, fill=USA_region)) +
geom_boxplot() +
labs(title="Log ppm by USA_region",
x="USA region",y="log(ppm)") + theme_classic() +
theme(legend.position="none",axis.text.x = element_text(angle = 90))

```
```{r}
#by state? hierarchical?
streetrx_morphine %>% count(state)
median(table(streetrx_morphine['state']))
table(streetrx_morphine['state']) #ignore count = 0 group,
# some groups does not have a lot of data
#varation by state?
#sample a few, 25?
```
```{r}
#varations among states
set.seed(1000)
sample_state <- sample(unique(streetrx_morphine$state),25,replace=F)
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,sample_state),],
aes(x=state, y=logppm,, fill=state)) +
geom_boxplot() +
labs(title="Log ppm by state",
x="state",y="log(ppm)") + theme_classic() +
theme(legend.position="none",axis.text.x = element_text(angle = 90))
#yes??, but some only have <50

sample_state <- which(table(streetrx_morphine$state)> 54) #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=state, y=logppm,, fill=state)) +
geom_boxplot() +
labs(title="Log ppm by state with >54 entries",
x="state",y="log(ppm)") + theme_classic() +
theme(legend.position="none",axis.text.x = element_text(angle = 90))
#do we need to do an anova test to see if grouped averages are truely different?

```
```{r}
streetrx_morphine %>% count(source_fac)
```
```{r}
#categorical variables:
#source, bulk purchase, USA region,
# we have seen that USA region does not matter??
# source
ggplot(streetrx_morphine,aes(x=source_fac, y=logppm, fill=source_fac)) +
geom_boxplot() + scale_fill_brewer(palette="Set1") +
labs(title="Log(ppm) vs Source", x="source",y="Log(ppm)") +
theme_classic() + theme(legend.position="none")
#can't say the difference

# state and source
sample_state <- sample(unique(streetrx_morphine$state),8,replace=F)
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,sample_state),],
aes(x=source_fac, y=logppm, fill=source_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs source by state",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#lets only do it for state with >50

sample_state <- which(table(streetrx_morphine$state)> 54)[1:10] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=source_fac, y=logppm, fill=source_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs source by state with >50 first 10/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[11:20] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=source_fac, y=logppm, fill=source_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs source by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[21:28] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=source_fac, y=logppm, fill=source_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs source by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
# slope on source_fac varies with state? yes? we can try comparing AIC values with and without
```

```{r}
# state and bulk purchase?
streetrx_morphine %>% count(bulk_fac)

#bulk
ggplot(streetrx_morphine,aes(x=bulk_fac, y=logppm, fill=bulk_fac)) +
geom_boxplot() + scale_fill_brewer(palette="Greens") +
labs(title="Log(ppm) vs Bulk purchase", x="bulk purchase?",y="Log(ppm)") +
theme_classic() + theme(legend.position="none")

ggplot(streetrx_morphine,aes(x=bulk_fac, y=ppm, fill=bulk_fac)) +
geom_boxplot() + scale_fill_brewer(palette="Greens") +
labs(title="ppm vs Bulk purchase", x="bulk purchase?",y="ppm") +
theme_classic() + theme(legend.position="none")

sample_state <- sample(unique(streetrx_morphine$state),8,replace=F)
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,sample_state),],aes(x=bulk_fac, y=logppm, fill=bulk_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs bulk purchase? by state",
x="bulk purchase?",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#lets only do it for state with >50

sample_state <- which(table(streetrx_morphine$state)> 54)[1:10] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=bulk_fac, y=logppm, fill=bulk_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs bulk purchase? by state with >50 first 10/28",
x="bulk purchase?",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[11:20] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=bulk_fac, y=logppm, fill=bulk_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs bulk purchase? by state with >50 11-20/28",
x="bulk purchase?",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[21:28] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=bulk_fac, y=logppm, fill=bulk_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs bulk purchase? by state with >50 11-20/28",
x="bulk purchase?",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#seems like slope on bulk_fac varies with state?
```

```{r}
# state and form
ggplot(streetrx_morphine,aes(x=form_fac, y=logppm, fill=form_fac)) +
geom_boxplot() + scale_fill_brewer(palette="Greens") +
labs(title="Log(ppm) vs form", x="form",y="Log(ppm)") +
theme_classic() + theme(legend.position="none")

sample_state <- sample(unique(streetrx_morphine$state),8,replace=F)
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,sample_state),],
aes(x=form_fac, y=logppm, fill=form_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs form by state",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#lets only do it for state with >50

sample_state <- which(table(streetrx_morphine$state)> 54)[1:10] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=form_fac, y=logppm, fill=form_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs form by state with >50 first 10/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[11:20] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=form_fac, y=logppm, fill=form_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs form by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[21:28] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=form_fac, y=logppm, fill=form_fac)) +
geom_boxplot() +
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs form by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#varting slope on form by state? we are not using them
```
```{r}
streetrx_morphine %>% count(form_fac)
#since only 3 out of XXX, and 
```
```{r}
#numerical: mgstr
ggplot(streetrx_morphine,aes(x=mgstr,y=logppm))+geom_point()+geom_smooth(method="lm",col="red3")
# mgstr shoud be included in the final model

sample_state <- which(table(streetrx_morphine$state)> 54)[1:10] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=mgstr, y=logppm, fill=form_fac)) +
geom_point() +geom_smooth(method="lm",col="red3")+
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs mgstr by state with >50 first 10/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[11:20] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
       aes(x=mgstr,y=logppm, fill=form_fac))+
geom_point() +geom_smooth(method="lm",col="red3")+
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs mgstr by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=5)

sample_state <- which(table(streetrx_morphine$state)> 54)[21:28] #since median is 54
ggplot(streetrx_morphine[is.element(streetrx_morphine$state,names(sample_state)),],
aes(x=mgstr, y=logppm, fill=form_fac)) +
geom_point() +geom_smooth(method="lm",col="red3")+
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs mgstr by state with >50 11-20/28",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ state,ncol=4)
#varying slope on mgstr across state?
```

```{r}
#interaction between variables:
ggplot(streetrx_morphine,
aes(x=mgstr, y=logppm, fill=bulk_fac)) +
geom_point() +geom_smooth(method="lm",col="red3")+
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs mgstr by bulk_fac",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ bulk_fac,ncol=2)

ggplot(streetrx_morphine,
aes(x=mgstr, y=logppm, fill=source_fac)) +
geom_point() +geom_smooth(method="lm",col="red3")+
scale_fill_brewer(palette="Set1") +
labs(title="log(ppm) vs mgstr by source_fac",
x="source",y="log(ppm") +
theme_classic() + theme(legend.position="none") +
facet_wrap( ~ source_fac,ncol=5)

ggplot(streetrx_morphine,aes(x=bulk_fac, y=logppm, fill=form_fac)) +
geom_boxplot() + scale_fill_brewer(palette="Greens") +
labs(title="Log(ppm) vs bulk by source_fac", x="form",y="Log(ppm)") +
theme_classic() + theme(legend.position="none")+
  facet_wrap(~source_fac)
```
```{r}

baseline <- lm(logppm~mgstr_c+source_fac+form_fac+bulk_fac,data=streetrx_morphine)
summary(baseline)
baseline$coefficients
plot(baseline)
plot(baseline,which=4)
plot(streetrx_morphine$mgstr_c,baseline$residuals)
# checking linearity? The distribution of residuals does not appear to have a trend. Linearity: checked 
```


```{r}
nullModel<- lm(logppm~1,data= streetrx_morphine)
fullModel <- lm(logppm~mgstr_c+source_fac+form_fac+bulk_fac+mgstr_c:source_fac+mgstr_c:bulk_fac+source_fac:bulk_fac,data=streetrx_morphine)

stepwise <- step(nullModel,scope = formula(fullModel),direction="both",trace=0)
stepwise$call
```
```{r}
# our EDA suggests bulk_fac might not be a significant factor, let's to a Anova F test to confirm if bulk_fac is really needed
model1 <- lm(formula = logppm ~ mgstr_c + bulk_fac + source_fac, data = streetrx_morphine)
model2 <- lm(formula = logppm ~ mgstr_c +  source_fac, data = streetrx_morphine)
anova(model1,model2)#anova F test

# check if sourc_fac is really significant
model3 <- lm(formula = logppm ~ mgstr_c +  bulk_fac, data = streetrx_morphine)
anova(model1,model3)


```

```{r}
# model suggested, no form_fac and no interaction term
model1 <- lm(formula = logppm ~ mgstr_c + bulk_fac + source_fac, data = streetrx_morphine)
summary(model1)
plot(model1)
plot(model1,which=4)
plot(model1,which=5)
#check no need to check VIC since only one is numerical

```
#interesting cluster: fitted values<-2.0


```{r}
streetrx_morphine$logppm_pred <- predict(model1)
a<-streetrx_morphine%>%filter(logppm_pred< -2.0)
# ggplot(a,aes(x=state))+geom_bar()+coord_flip()+labs(title="Predicted log(ppm) < -0.2")
b<-streetrx_morphine%>%filter(logppm_pred>= -2.0)
hist(a$mgstr_c)
hist(b$mgstr_c)


# ggplot(b,aes(x=state))+geom_bar()+coord_flip()+labs(title="Predicted log(ppm) >= -0.2")

```

```{r}
# try hierarchical with state
hier_model1 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(1|state),data=streetrx_morphine)
summary(hier_model1)


# hier_model2 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(1|state/USA_region),data=streetrx_morphine)
# summary(hier_model2)

hier_model2 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(1|state)+
                       (1|USA_region),data=streetrx_morphine)

summary(hier_model2)


AIC(hier_model1)
AIC(hier_model2) #not better, not by region
anova(hier_model1,hier_model2)

```
```{r}
hier_model3 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(mgstr_c|state),data=streetrx_morphine)
summary(hier_model3)
anova(hier_model1,hier_model3) 
AIC(hier_model1)
AIC(hier_model3)
# AIC much smaller, p<0.05 yes!! varying slope on mgstr_c
# across states, the slope on mgstr_c is not significantly different (std = 0.088)

# what exactly explains residual std of 0.90? variation across states is too small compared to the variations among each observation. std =0.90.

```

```{r}
plot(hier_model3)
plot(streetrx_morphine$mgstr_c,residuals(hier_model3))
qqnorm(residuals(hier_model3))
```
```{r}
hier_model4 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(mgstr_c+bulk_fac|state),data=streetrx_morphine)
summary(hier_model4)
AIC(hier_model4)
anova(hier_model3,hier_model4) 

## the warning means the model is overfitted
```

```{r}
hier_model5 <- lmer(logppm~mgstr_c+bulk_fac+source_fac+(mgstr_c+source_fac|state),data=streetrx_morphine)
summary(hier_model5)
AIC(hier_model5)
anova(hier_model3,hier_model5) 
```


```{r}
finalmodel <- hier_model3
summary(finalmodel)
#confint(finalmodel)
dotplot(ranef(finalmodel,condVar=TRUE))$state

```
```{r}
# how do we make senese f these confidence intervals?
# what is .sig02, sig03,
confint(finalmodel)

```
```{r}
# outliers
```



```{r}
exp(fixef(finalmodel))
# coef(summary(finalmodel))
```


```{r}

rand_effects <- ranef(finalmodel,condVar=TRUE)$state
rand_effects[c("Massachusetts","Michigan","North Carolina"),]
basline <-exp(fixef(finalmodel)[1]+rand_intercept[c("Massachusetts","Michigan","North Carolina"),"(Intercept)"])
mgstr_effects <- exp(fixef(finalmodel)[2]+rand_effects[c("Massachusetts","Michigan","North Carolina"),"mgstr_c"])

data.frame(state =c("Massachusetts","Michigan","North Carolina"), basline=basline,mgstr_effects=mgstr_effects)
```
Average ppm of morphine in the U.S. is 0.40.

For any morphine purchased in MA, source = Heard it, bulk =No, mgstr_c = avg, the baseline ppm (price per milligram) is 0.47, higher than the national average price

The baseline ppm in NC is 0.31

For any morphine purchased in MA, a 1 unit increase in mgstr reduces the ppm by a multiplicative effect of 0.78, that is about a 0.22 reduction, lower than the average effect of mgstr_c in the U.S.





